---
title: "SCohenDevries_Final"
author: "Sam Cohen-Devries"
date: "5/27/2019"
output:
  word_document: default
  html_document: default
---
*Generate a random variable X that has 10,000 random uniform numbers from 1 to N*

```{r}
N <- 8
mu <- (N+1)/2
sig <- mu
X <- runif(10000,0,N)
```

*Then generate a random variable Y that has 10,000 random normal numbers with a mean of mu=sigma=(N+1)/2.*

```{r}
Y <- rnorm(10000,mu,sig)
```

*"x" is estimated as the median of the X variable*

```{r}
x <- median(X)
```

*"y" is estimated as the 1st quartile of the Y variable*

```{r}
y <- quantile(Y,0.25)

pX_gt_x <- length(X[X>x])/length(X)
pX_gt_y <- length(X[X>y])/length(X)

#P(X>x|X>y) = P(X>x intersect X>y)/P(X>y) = P(X>x)/P(X>y)
pX_gt_x/pX_gt_y

#P(X>x, Y>y)
pX_gt_x*pX_gt_y

#P(X<x | X>y)
length(X[X<x&X>y])/length(X[X>y])
```

*Investigate whether P(X>x and Y>y)=P(X>x)P(Y>y) by building a table and evaluating the marginal and joint probabilities.*

```{r}
df <- data.frame(cbind(X,Y))
df$y <- y
df$x <- x
df2 <-subset(df,X>x & Y>y)
nrow(df2)/nrow(df)
(nrow(subset(df,X>x))/nrow(df)) * (nrow(subset(df,Y>y)))/nrow(df)
```

*Check to see if independence holds by using Fisher’s Exact Test and the Chi Square Test.*

```{r}
#fisher.test(X,Y)
chisq.test(table(X,Y))
```

*What is the difference between the two? Which is most appropriate?*
Fisher's exact test is most accurate when working with small sample sizes, which is not the case here. Fisher's tests also provides an exact p-value, but it makes some assumptions about the data structure that may or may not be accurate. For larger sample sizes, we can expect the Chi-Square to be most useful. In this case, the Chi Square Test appears to be most appropriate.

---------------------------------------------------------------------------------------------------------------------------------------------------

*https://www.kaggle.com/c/house-prices-advanced-regression-techniques*
```{r,echo=FALSE}
library(corpcor)
library(matrixcalc)
library(e1071)
library(MASS)

train <- read.csv('/Users/samandleo/Downloads/train.csv')
```

*Provide univariate descriptive statistics and appropriate plots for the training data set*

```{r}
summary(train)

hist(train$SalePrice)
plot(SalePrice ~ GrLivArea,data=train)
```

*Provide a scatterplot matrix for at least two of the independent variables and the dependent variable*

```{r}
sapply(train,class) #checking for numeric variables
res <- cor(train[,c("SalePrice","OpenPorchSF","YearRemodAdd")])
round(res, 2)
```

*Derive a correlation matrix for any three quantitative variables in the dataset*

```{r}
cm <- pairs(~SalePrice+LotArea+OverallQual,data=train)
```

*Test the hypotheses that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval.*

```{r}
cor.test(train$SalePrice,train$OverallQual,conf.level=.8)
cor.test(train$SalePrice,train$LotArea,conf.level=.8)
cor.test(train$OverallQual,train$SalePrice,conf.level=.8)
cor.test(train$OverallQual,train$LotArea,conf.level=.8)
cor.test(train$LotArea,train$SalePrice,conf.level=.8)
cor.test(train$LotArea,train$OverallQual,conf.level=.8)
```

*Discuss the meaning of your analysis.  Would you be worried about familywise error? Why or why not?*

Using the cor.test function in R, we were able to reject the null hypothesis that the correlation is 0, with a high signifigance (p-value < 0.05). Due to the high level of signifigance, I would not be worried about familywise error.

*Invert your correlation matrix from above. *

```{r}
pm <- solve(res)
```

*Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix. *

```{r}
pm %*% res
res %*% pm
```

*Conduct LU decomposition on the matrix.*

```{r}
lu.decomposition(pm)
```

*Select a variable in the Kaggle.com training dataset that is skewed to the right, shift it so that the minimum value is absolutely above zero if necessary*

```{r}
skews <- data.frame(colName = character(), skewness = numeric())
for(i in colnames(train)){
    if(class(train[,i])=="integer" & !is.na(skewness(train[,i]))){
    skews <- rbind(skews,data.frame(i,skewness(train[,i])))
    }
}
skews

train$LotArea.trans <- train$LotArea ^ 2
```

*fit an exponential probability density function*

```{r}
lot.fit <- fitdistr(train$LotArea.trans, densfun="exponential")
lot.fit.func <- function(x){lot.fit$estimate * exp(-lot.fit$estimate * x)}
```

*Find the optimal value of lamba for this distribution, and then take 1000 samples from this exponential distribution using this value *

```{r}
opt.lam <- optim(lot.fit$estimate,lot.fit.func)
opt.samp <- rexp(1000,opt.lam$value)
```

*Plot a histogram and compare it with a histogram of your original variable*

```{r}
hist(opt.samp)
hist(train$LotArea)
```

*Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF)*

```{r}
qexp(.05,rate = opt.lam$value)
qexp(.05,rate = opt.lam$value, lower.tail = FALSE )
```

*generate a 95% confidence interval from the empirical data, assuming normality*

```{r}
ci <- qt(1-(.95/2),df=length(train$LotArea)-1)*sd(train$LotArea)/sqrt(length(train$LotArea))
print(cbind(mean(train$LotArea)-ci,mean(train$LotArea)+ci))
```

*provide the empirical 5th percentile and 95th percentile of the data*

```{r}
quantile(train$LotArea,.05)
quantile(train$LotArea,.95)
```

*Build some type of multiple regression  model*

```{r}
colSums(is.na(train))
drops <- c("MiscFeature","Fence","PoolQC","FireplaceQu","Alley")
train<- train[ , !(names(train) %in% drops)]

for(i in colnames(train)){
    train[,i][is.na(train[,i])] <- sample(train[,i][!is.na(train[,i])],length(train[,i][is.na(train[,i])]))
}

train.lm <- lm(SalePrice ~ ., data = na.omit(train))
#train.lm.stepped <- step(train.lm, direction = "backward", trace=FALSE ) 
#summary(train.lm.stepped)

#keeping only columns w high signifigance; eliminating redundant variables (BsmtSF,GarageCond...)
keeps <- c("LotArea"
,"LandSlope"
,"Neighborhood"
,"Condition1"
,"Condition2"
,"OverallQual"
,"OverallCond"
,"YearBuilt"
,"RoofMatl"
,"ExterQual"
,"BsmtQual"
,"TotalBsmtSF"
,"GarageQual"
,"PoolArea"
,"SaleCondition"
,"SalePrice"
)
train2 <- train[,keeps]
train.lm2 <- lm(SalePrice ~ ., data = na.omit(train2))
summary(train.lm2)

drops2 <- c("Condition1","Condition2","YearBuilt","GarageQual","SaleCondition")
train3 <- train2[ , !(names(train2) %in% drops2)]

train.lm3 <- lm(SalePrice ~ ., data = na.omit(train3))
summary(train.lm3)

plot(fitted(train.lm3),resid(train.lm3))
qqnorm(resid(train.lm3))
qqline(resid(train.lm3))

drops3 <- c("Neighborhood")
train4 <- train3[ , !(names(train3) %in% drops3)]

train.lm4 <- lm(SalePrice ~ ., data = na.omit(train4))
summary(train.lm4)

test <- read.csv('/Users/samandleo/Downloads/test.csv')
keeps <- keeps[keeps != "SalePrice"]
test <- test[,c(keeps,"Id")]
test <- test[ , !(names(test) %in% drops2)]
for(i in colnames(test)){
    test[,i][is.na(test[,i])] <- sample(test[,i][!is.na(test[,i])],length(test[,i][is.na(test[,i])]))
}

test$SalePrice <- predict(train.lm3,test)
colnames(test)
submission.scd <- subset(test,select=c("Id","SalePrice"))

write.csv(submission.scd,file="submission_scd.csv",row.names=FALSE)
```

#Username: SamCD
#Score: 0.20441